### 2025-05-29 docs/article/Sections/problem-statement.tex
Highlight Reactive versus Suspend for HtmlFlow templates and cross ref thos Listings in Results analysis.

\begin{center}
\begin{minipage}{0.50\textwidth}
\begin{lstlisting}[
    language=Kotlin,
    basicstyle=\scriptsize\ttfamily,
    numbers=none,
    caption={\textit{HtmlFlow} presentation template in Koltin with an \texttt{Observable} model},
    label={lst:presentation-observable}
]
await { div, model, onCompletion ->
  model
    .doOnNext { presentation ->
      presentationFragmentAsync
        .renderAsync(presentation)
        .thenApply { frag -> div.raw(frag) }
      }
    .doOnComplete { onCompletion.finish() }
    .subscribe()
}
\end{lstlisting}
\end{minipage}
\hfill
\begin{minipage}{0.46\textwidth}
\begin{lstlisting}[
    language=Kotlin,
    basicstyle=\scriptsize\ttfamily,
    numbers=none,
    label={lst:presentation-flow},
    caption={\textit{HtmlFlow} presentation template in Kotlin with a \texttt{Flow} model},
]
suspending { model ->
  model
    .toFlowable()
    .asFlow()
    .collect { presentation ->
      presentationFragmentAsync
        .renderAsync(presentation)
        .thenApply { frag -> raw(frag) }
    }
}
\end{lstlisting}
\end{minipage}
\end{center}

>>>>

\begin{center}
\begin{minipage}{0.50\textwidth}
\begin{lstlisting}[
    language=Kotlin,
    basicstyle=\scriptsize\ttfamily,
    numbers=none,
    caption={\textit{HtmlFlow reactive} presentation template in Koltin with an \texttt{Observable} model},
    label={lst:presentation-observable}
]
await { div, model, onCompletion ->
  model
    .doOnNext { presentation ->
      presentationFragmentAsync
        .renderAsync(presentation)
        .thenApply { frag -> div.raw(frag) }
      }
    .doOnComplete { onCompletion.finish() }
    .subscribe()
}
\end{lstlisting}
\end{minipage}
\hfill
\begin{minipage}{0.46\textwidth}
\begin{lstlisting}[
    language=Kotlin,
    basicstyle=\scriptsize\ttfamily,
    numbers=none,
    label={lst:presentation-flow},
    caption={\textit{HtmlFlow suspend} presentation template in Kotlin with a \texttt{Flow} model},
]
suspending { model ->
  model
    .toFlowable()
    .asFlow()
    .collect { presentation ->
      presentationFragmentAsync
        .renderAsync(presentation)
        .thenApply { frag -> raw(frag) }
    }
}
\end{lstlisting}
\end{minipage}
\end{center}

### 2025-05-29 docs/article/Sections/results.tex
Highlight Reactive versus Suspend for HtmlFlow templates and cross ref thos Listings in Results analysis.

The results in Figure~\ref{fig:presentations-webflux-jmeter} depict the
throughput (number of requests per second) for each template engine, with
concurrent users ranging from 1 to 128, from left to right. The benchmarks
include HtmlFlow using suspendable web templates (HtmlFlow-Susp), Jstachio
using Virtual Threads with the \texttt{Iterable} interface (Jstachio-Virtual),
and Thymeleaf using the reactive View Resolver driver (Thymeleaf-Rx).
\textit{Blocking} and \textit{Virtual} represent the average throughput of the
blocking approaches (i.e., KotlinX, Rocker, Jstachio, Pebble, Freemarker,
Trimou, HtmlFlow, and Thymeleaf) when run in the context of a separate
coroutine dispatcher or Virtual Threads, respectively.

We show the HtmlFlow-Susp, Jstachio-Virtual, and Thymeleaf-Rx engines
separately to observe the performance of the non-blocking engines when using
the Suspending, Virtual Thread, and Reactive approaches. The \textit{Blocking}
and \textit{Virtual} are aggregated due to the similar performance of different
engines when using those approaches.

@@

The results for the Spring MVC implementation, shown in
Figure~\ref{fig:presentations-springmvc-jmeter}, compare two synchronous
approaches: \textit{Blocking}, which uses platform threads with
\texttt{StreamingResponseBody}, and \textit{Virtual}, which uses Virtual
Threads. Note that for Spring MVC, which follows a thread-per-request
architecture, neither of the asynchronous approaches—\textbf{Reactive} nor
\textbf{Suspendable}—described in Section~5 are applicable. Regarding the
synchronous strategy, both \textit{Blocking} and \textit{Virtual} scale
effectively up to 128 concurrent users, with the Virtual Threads approach
achieving a slightly higher throughput of approximately 9,000 requests per
second. However, these values are slightly lower than those observed in the
Spring WebFlux implementation.

@@

Additionally, \textit{HtmlFlow-Rx}, a reactive implementation of the HtmlFlow
template engine that utilizes Quarkus's reactive programming model, achieved
throughput similar to the \textit{Blocking} and \textit{Virtual}
approaches—approximately 10,000 requests per second. This demonstrates that
Quarkus's reactive programming model is effective for enabling PSSR\@. However,
there is no significant difference in performance between the reactive and
synchronous approaches, indicating that the reactive model does not provide a
substantial advantage over synchronous approaches in this context.

>>>>

The results in Figure~\ref{fig:presentations-webflux-jmeter} depict the
throughput (number of requests per second) for each template engine, with
concurrent users ranging from 1 to 128, from left to right. The benchmarks
include HtmlFlow using suspendable web templates (\textit{HtmlFlow-Susp},
equivalent to the approach shown in Listing~\ref{lst:presentation-flow}),
Jstachio using Virtual Threads with the \texttt{Iterable} interface
(\textit{Jstachio-Virtual}), and Thymeleaf using the reactive View Resolver
driver (Thymeleaf-Rx).
\textit{Blocking} and \textit{Virtual} represent the average throughput of the
blocking approaches (i.e., KotlinX, Rocker, Jstachio, Pebble, Freemarker,
Trimou, HtmlFlow, and Thymeleaf) when run in the context of a separate
coroutine dispatcher or Virtual Threads, respectively.

We show the \textit{HtmlFlow-Susp}, \textit{Jstachio-Virtual}, and \textit{Thymeleaf-Rx} engines
separately to observe the performance of the non-blocking engines when using
the Suspending, Virtual Thread, and Reactive approaches. The \textit{Blocking}
and \textit{Virtual} are aggregated due to the similar performance of different
engines when using those approaches.

@@

The results for the Spring MVC implementation, shown in
Figure~\ref{fig:presentations-springmvc-jmeter}, compare two synchronous
approaches: \textit{Blocking}, which uses platform threads with
\texttt{StreamingResponseBody}, and \textit{Virtual}, which uses Virtual
Threads. Note that for Spring MVC, which follows a thread-per-request
architecture, neither of the asynchronous approaches—\textbf{Reactive} nor
\textbf{Suspendable}—described in Section~\ref{sec:bench} are applicable. Regarding the
synchronous strategy, both \textit{Blocking} and \textit{Virtual} scale
effectively up to 128 concurrent users, with the Virtual Threads approach
achieving a slightly higher throughput of approximately 9,000 requests per
second. However, these values are slightly lower than those observed in the
Spring WebFlux implementation.

@@

Additionally, \textit{HtmlFlow-Rx}, a reactive implementation of the HtmlFlow
template engine (equivalent to the approach shown in
Listing~\ref{lst:presentation-observable}) that utilizes Quarkus's reactive
programming model, achieved throughput similar to the \textit{Blocking} and
\textit{Virtual}
approaches—approximately 10,000 requests per second. This demonstrates that
Quarkus's reactive programming model is effective for enabling PSSR\@. However,
there is no significant difference in performance between the reactive and
synchronous approaches, indicating that the reactive model does not provide a
substantial advantage over synchronous approaches in this context.
